

# INSTRUCTIONS

-   This is an individual assignment.
-   Submit your answer digitally as two files through Moodle:
    -   An R markdown file (extension **Rmd**). Use the template provided
        to you and provide your answers (both code and text) below each
        question.
    -   An **HTML** file "knitted" by RStudio including all the results and
        plots. More details on how to create these files will be provided
        in class on week 3.
-   Follow the Style Guide (available on Moodle). You can be
    penalized on up to 20% in each question for which you do not follow
    the Style Guide.
-   Questions regarding the assignment should be posted <span class="underline">exclusively</span> on
    the respective discussion forum on Moodle.


## <span class="underline">Deadline:</span> Monday, February 21 at 23:59.

-   Late submissions are <span class="underline">not allowed</span>

<span class="underline">**Warning:**</span> The detection of <span class="underline">any form of plagiarism</span> in your work
means the assignment will be graded with <span class="underline">ZERO points</span>.

\newpage


# Movie Networks

We are interested in assessing what are the most important movies in
the decade 2010-2019. We will use different strategies to do so.
First, we will load and prepare the data.


## Load and prepare the data

The first step is to load and prepare the movie data. The following
instructions perform some routine data preparation operations. Each
set of instructions is preceded by a comment explaining the procedure.
Run the code below and try to understand each line of code as you
might need to perform some changes.

```{r   }
library(data.table)     # Run once per session
library(ggplot2)        # Run once per session

# Load data from file 20200120-imdb_movie_actor.csv (do not forget to
# change your working directory to the folder containing the
# 20200120-imdb_movie_actor.csv file)
dt.movie.actor <- fread("20200120-imdb_movie_actor.csv") 

# Count in how many movies each actor has participated and how many 
# principal actor each movie has
dt.movie.actor[, n_movies := .N, by=actor]
dt.movie.actor[, n_actors := .N, by=list(movie, year)]

# Remove entries in which actors have no name 
dt.movie.actor <- dt.movie.actor[!(actor == "")]

# Save dt.movie.actor. Next time you can simply call the load function (below)
save(dt.movie.actor, file="imdb_movie_actor.RData") 
```

Load the data that you prepared using the instructions below. As mentioned in the
comments, you can start from this line if you have previously saved these data.

```{r   }
# Load previously saved dt.movie.actor. You can
# start in this line if you have previously saved these data.
load("imdb_movie_actor.RData") 
```


## Questions (`data.table`) `[7 points]`

This set of questions require that you know how to manipulate a
`data.table`. Answer each of the following questions below. Include
all the code you created/used in your answer.

1.  What is the total amount of movies in the `dt.movie.actor` dataset?
    `[1 point]`

```{r}
length(unique(dt.movie.actor[, movie]))
# The dt.movie.actor dataset has a total amount of 368522 movies.
```

2.  List the actors from the movie `"Fight Club (1999)"`. List
    the actors from the movie `"Se7en (1995)"`. `[1 point]`

```{r}
dt.movie.actor[movie == "Fight Club (1999)", list(movie, actor)] 
dt.movie.actor[movie == "Se7en (1995)", list(movie, actor)] 
#The actors from 'Fight Club (1999)' are Brad Pitt (b.1963), Meat Loaf (b.1947), Edward Norton (b.1969) and Zach Grenier (b.1954).
#The actors from 'Se7en (1995)' are Brad Pitt (b.1963), Morgan Freeman (b.1937), Kevin Spacey (b.1959) and Andrew Kevin Walker (b.1964).
```

3.  Which actors participated on both movies? Hint: The function
    `intersect` calculates the intersection of two sets. `[1 point]`

```{r}
intersect(dt.movie.actor[movie == "Fight Club (1999)", actor],
          dt.movie.actor[movie == "Se7en (1995)", actor])
# The only actor that has participated in both movies is Brad Pitt (b.1963).
```

4.  In which movies did Brad Pitt (b.1963) and George Clooney (b.1961) star together?
    `[1 point]`

```{r}
intersect(dt.movie.actor[actor == "Brad Pitt (b.1963)", movie],
          dt.movie.actor[actor == "George Clooney (b.1961)", movie])
#Brad Pitt and George Clooney star together in 'Ocean's Eleven (2001)', 'Ocean's Twelve (2004)', 'Ocean's Thirteen (2007)' and 'Burn After Reading (2008)'.
```

5.  Create a table that shows the number of movies released per year.
    This table should include three columns: `year`, `n_movies`, and
    `csum_n_movies`. The first column should contain the year, the
    second the number of movies in that year, and the third, the number
    of movies released since the first year in the data and up to the
    year in that line. Tip: Use the function `cumsum` and check if the
    amount in the last year is the same as the total number of movies
    in question 1. `[1 point]`

```{r}
movies.per.year <- unique(dt.movie.actor, by="movie")[, .(n_movies = .N), by=year]
movies.per.year.ascending <- movies.per.year[order(year)]
movies.per.year.ascending[, csum_n_movies := cumsum(n_movies)]
last(movies.per.year.ascending)


```

6.  Which actor/actress has starred in the most movies across all data?
    After (and including) 2000, which year has the most movie
    participations by a single actor/actress? Who is that
    actor/actress? What do these two actors/actresses have in common?
    `[1 point]`

```{r}
unique(dt.movie.actor, by="actor")[order(-n_movies)][1]$actor
dt.movie.actor[dt.movie.actor$year >= 2000][, .(n_movies=.N), by=list(year,actor)][order(-n_movies)][1]$year
dt.movie.actor[dt.movie.actor$year >= 2000][, .(n_movies=.N), by=list(year,actor)][order(-n_movies)][1]$actor
# Through a quick Google Search I can see that both actors are from India, which makes sense, as the Indian movie industry is very developed.
```

7.  Consider only the 10% most popular movies (by votes) in the decade 2010-2019.
    List the top 10 actors that starred in the most movies in the decade.
    Which year(s) has/have the most movie participations by a single actor?
    Hint: you can use the function `quantile` to find how many votes
    does the movie in percentile 90 have. `[1 point]`


```{r}
top.movies <- dt.movie.actor[dt.movie.actor$votes >= quantile(dt.movie.actor$votes, probs=0.9, na.rm=TRUE) & dt.movie.actor$year >= 2010 & dt.movie.actor$year <= 2019]
top.movies[, n_movies := .N, by = actor]
unique(top.movies, by="actor")[order(-n_movies)][1:10]$actor
head(top.movies[, .N, by=list(actor,year)][order(-N)])
# The top 10 actors that starred in the most movies in the 2010-2019 decade were "Akshay Kumar (b.1967)", "Nicolas Cage (b.1964)", "James Franco (b.1978)", "John Cusack (b.1966)", "Robert De Niro (b.1943)", "Bruce Willis (b.1955)", "Anna Kendrick (b.1985)", "Samantha Ruth Prabhu (b.1987)", "Samuel L. Jackson (b.1948)", and "Prakash Raj (b.1965)"
# The years that had the most movie participations by a single actor were 2012, 2014, and 2015, with 7 movie participations  by Christian Slater (b.1969), Liam Neeson (b.1952), and James Franco (b.1978)
```

## Questions (`ggplot2`) `[3 points]`

1.  Plot a histogram with the number of movies per year. Which patterns
    do you observe? Is there anything strange? `[1 point]`

```{r}
#ggplot(dt.movie.actor, aes(x = year, y = n_movies)) + geom_histogram()
ggplot(dt.movie.actor) + geom_histogram(aes(x=year), binwidth=1) + xlim(c(1900,2025))
# From the plot, it is clear that the movie industry grew exponentially over the time-period. We see that the amount of movies made per year after the year 2000 skyrocketed. I don't feel like it is a strange behaviour, but actually the contrary. As technology evolved, so did the movie industry, and its demand. With better technology in film-making, it was expected that the number of movies made per year would have an exponential behavior.

```

2.  Plot a histogram that represents the distribution of number of IMDb
    votes per movie. The x-axis should represent the number of votes
    and the y-axis should represent how many movies have x number of
    votes. Which patterns do you observe? `[1 point]`

```{r}
ggplot(unique(dt.movie.actor, by="movie"), aes(x=votes)) + geom_histogram(bins=150) + xlim(c(0,11000)) + ylim(c(0,45000))
# The distribution of the IMDB votes seems to be skewed to the left, which indicates that most movies have a small number of votes, bringing the average votes per movie down. Only very few, selected, movies have many votes.
```

3.  Plot a histogram that represents the distribution of the number of
    actors per movie. The x-axis should represent the number of actors
    and the y-axis should represent how many movies have x number of actors.
    Describe your findings. `[1 point]`

```{r}
ggplot(unique(dt.movie.actor, by="movie"), aes(x=n_actors)) + geom_histogram(binwidth=1) + xlim(c(0,15))
#As is notorious, most movies have 4 actors in our data. Beside that, there is relatively the same number of movies with 1 to 8 actors (excluding 4), and only a few selected movies have more than 8 actors.
```

## Questions (`igraph`) `[10 points]`

1.  From this question onwards, and until the end of the assignment,
    focus only on <span class="underline">the actors that participated on the top 50 most
    popular movies from the 2010-2019 decade (by number of votes).</span>
    Load the `igraph` package and create a bipartite graph in which the
    edges correspond to actors&rsquo; participation in movies. How many movie
    participations exist? `[1 point]`
    
```{r}
library(igraph)

top.movies <- dt.movie.actor[year >= 2010 & year <= 2019, .N,
                             by = list(movie, votes)][order(-votes)][1:50]
dt.top.actors.movies <- dt.movie.actor[movie %in% top.movies$movie]
top.movies.decade <- dt.top.actors.movies[, list(name=unique(movie), type=TRUE)]
top.actors.decade <- dt.top.actors.movies[, list(name=unique(actor), type=FALSE)]

all.vertices <- rbind(top.movies.decade,top.actors.decade)

g <- graph.data.frame(dt.top.actors.movies[, list(actor,movie)], directed=FALSE, vertices=all.vertices)

g.movies <- bipartite.projection(g)$proj2

length(E(g))
#There are 200 movie participations across the actors that participated in the top 50 most popular movies from the 2010-2019 decade.
```


2.  Create a graph in which two movies are connected to each other if
    they have <span class="underline">at least one actor in common</span>. Calculate the <span class="underline">degree
    centrality</span> for each of the movies, and remove movies with no
    connections to other movies. <span class="underline">Hint:</span> the function
    `induced.subgraph` allows the creation of graphs with only a subset
    of the vertices. Calculate the following additional centrality
    measures for each of these movies: `[2 points]`
    -   Closeness centrality
    -   Betweenness centrality
    -   Eigenvector centrality

```{r}
subgraph <- induced.subgraph(g.movies, vids = degree(g.movies)!=0)

degree.centrality <- degree(g.movies)
closeness.centrality <- closeness(g.movies)
betweenness.centrality <- betweenness(g.movies)
evcent.centrality <- evcent(g.movies)$vector


```

3.  For each centrality measure, list the top 20 movies with highest
    centrality. How do you interpret the outcomes? `[2 points]`

```{r}
dt.g.movies <- data.table(get.data.frame(g.movies, "vertices"))
head(dt.g.movies[order(-degree.centrality)], 20)
head(dt.g.movies[order(-closeness.centrality)], 20)
head(dt.g.movies[order(-betweenness.centrality)], 20)
head(dt.g.movies[order(-evcent.centrality)], 20)
# From these results, we see that 'Avengers: Age of Ultron (2015)' has the highest degree centrality, which means it is the movie in the network that has the most actors in common with other movies in the network. Then, 'Shutter Island (2010)' has the highest closeness centrality, which indicates it is closely connected to the other movies, thus sharing very similar pools of actors. 'Shutter Island (2010)' also has the highest betweenness centrality, which measures how well situated it is on the network. This movie has the shortest average path to other movies in the Network. Finally, 'Avengers: Endgame (2019)' has the highest eigenvector centrality, which in turn reveals the importance of this node. This measure of centrality shows that this movie is connected to the most important movies, which means it is connected to the movies with the highest average votes from IMDB.

```

4.  Calculate the average clustering coefficient for the movies
    network. `[1 point]`

```{r}
transitivity(g.movies, type="average")
# The higher the clustering coefficient, the more direct connections the movies have between each other. That is, the more actors they have in common. As seen, the clustering coefficient of this network is 0.6551873, which is high and to be expected in real world networks.
```

5.  Choose one movie you like and plot the movie, their direct
    neighbors and the links among them. What is the clustering
    coefficient of this movie? Which is the actor with most
    participations among these (neighbor) movies, but not having
    participated in the movie itself? `[2 points]`

```{r}
my.movie <- neighbors(g.movies, "Shutter Island (2010)")
my.movie.graph <- induced.subgraph(g.movies, vids = (V(g.movies)$name %in% my.movie$name) |
                                   (V(g.movies)$name == "Shutter Island (2010)"))
plot(my.movie.graph, vertex.size=25)

transitivity(my.movie.graph, type="local", vids="Shutter Island (2010)")

my.movie.actors <- dt.movie.actor[movie == "Shutter Island (2010)", list(actor)]
neighbor.movies <- data.table(V(my.movie.graph)$name)
neighbor.movies.actors <- dt.top.actors.movies[movie %in% neighbor.movies[,V1]]
actors.only.in.neighbor <- data.table(neighbor.movies.actors[!neighbor.movies.actors$actor %in% my.movie.actors$actor])
actor <- unique(dt.movie.actor[dt.movie.actor$actor %in% actors.only.in.neighbor[,actor]], by="actor")[order(-n_movies)][1]$actor
actor

#As can be seen in the plot, 'Shutter Island (2010)' has some connections, which makes sense given that it has rather famous actors. We can see that the cluster coefficient is 0.428713 which illustrates its connection to its neighbors through actors in common. Finally, the actor with most participations among the neighbor movies of 'Shutter Island (2010)', while not having participated in the movie itself is Woody Harrelson (b.1961)"
```

6.  Plot the degree distribution of the movies. How do you compare them
    with the degree distribution of a random graph? What can be
    plausible explanations for the observed differences? `[2 points]`


```{r}
set.seed(123)
g.random <- sample_gnm(50,100)

ggplot() + geom_histogram(aes(x = degree(g.random)), binwidth=1)

ggplot() + geom_histogram(aes(x=degree(g.movies)), binwidth = 1)

#As can be seen, the degree distribution of a random graph follows a normal distribution, as expected. However, the degree distribution of our network doesn't, which is to be expected in a real-life scenario. We see that the degree distribution of the movies is skewed to the left, which indicate a shift in the mean degree distribution. The result comes as no surprise as most movies do not share actors with eachother, while only some have the most famous actors in common.
```
